### 2025-01-13 paper/icml25/first/sects/ablation.tex
minor + figs

@@

In fact, the transition function is the most complex component to replace as it requires modeling with precision the observations obtained from the environment after a number of $H$ actions from each agent. Avoiding the need of such network lightens the algorithm but it reintroduce the need for a differentiable environment. Nonetheless, the need for a differentiable reward, made by SHAC, is still relaxed.

Studying such setup can provide insights on the importance and difficulty of the transition function in the SHAC algorithm. TODO

>>>>

\begin{figure*}[!t]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-ablation-1-mlp.pdf}
        \caption{Dispersion, MLP, 1 agent}
        \label{fig:dispersion-ablation-mlp-1}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-ablation-3-transformer.pdf}
        \caption{Dispersion, Transformer, 3 agents}
        \label{fig:dispersion-ablation-transformer-3}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-ablation-5-transformer.pdf}
        \caption{Dispersion, Transformer, 5 agents}
        \label{fig:dispersion-ablation-transformer-5}
    \end{subfigure}

    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/transport-ablation-1-mlp.pdf}
        \caption{Transport, MLP, 1 agent}
        \label{fig:transport-ablation-mlp-1}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/transport-ablation-3-transformer.pdf}
        \caption{Transport, Transformer, 3 agents}
        \label{fig:transport-ablation-transformer-3}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/transport-ablation-5-transformer.pdf}
        \caption{Transport, Transformer, 5 agents}
        \label{fig:transport-ablation-transformer-5}
    \end{subfigure}

    \caption{Comparison between \fname{} with and without transition network, labelled \fnamer{}, for increasing number of agents for Dispersion, Transport, Discovery, and Sampling scenarios.}
    \label{fig:ablation}
\end{figure*}

@@

In fact, the transition function is the most complex component to replace as it requires modeling with precision the observations obtained from the environment after a number of $H$ actions from each agent. Avoiding the need of such network lightens the algorithm but it reintroduce the need for a differentiable environment.

Fig.~\ref{fig:ablation} summarizes the results. Most notably, using the true transition function does not appear to provide any significant advantage in terms of reward achieved. In fact, in some cases, we report better results for \fname{}, namely in Fig.~\ref{fig:transport-ablation-1-mlp}.

### 2025-01-13 paper/icml25/first/sects/background.tex
minor + figs

\section{Background}

\subsection{Markov Decision Processes}
To provide context for our work, we first introduce the reinforcement learning (RL) framework. Each task is modeled as a finite-horizon Markov Decision Process (MDP) $\mathcal{M} = (\mathcal{S}, \mathcal{A}, \hat{F}, \hat{R}, \gamma, T)$, where $\mathcal{S}$ represents the state space, $\mathcal{A}$ denotes the action space, $\hat{F}:\mathcal{S} \times \mathcal{A} \rightarrow \mathcal{S}$ is the transition function, $\hat{R}:\mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow \mathbb{R}$ is the reward function, $\gamma \in [0, 1]$ is the discount factor, and $T \in \mathbb{N}$ is the time horizon.

Further, let $\mathcal{T}$ denote the space of trajectories, defined as $\mathcal{T} = (\mathcal{S} \times \mathcal{A} \times \mathcal{S} \times \mathbb{R})^*$, where each trajectory consists of a sequence of starting states, actions, subsequent states, and rewards. We will use the symbols $((s_0, a_0, s_1, r_0), (s_1, a_1, s_2, r_1), \ldots)$ to denote a trajectory in $\mathcal{T}$.

The goal of the MDP is to find a policy $\pi:\mathcal{S}\rightarrow\mathcal{A}$ that maximizes the expected return 
$$J(\pi) = \mathbb{E}_{\tau\sim\pi}\left[\sum_{t=0}^{T-1}\gamma^t r(s_t, a_t, s_{t+1})\right]$$
Where $\tau \in \mathcal{T}$ is a trajectory generated by the policy $\pi$ with the transition function $\hat{F}$.

@@

To solve an MDP $\mathcal{M} = (\mathcal{S}, \mathcal{A}, \hat{F}, \hat{R}, \gamma, T)$, SHAC trains a policy $\pi$ by minimizing the loss function $\mathcal{L}_\theta^{\pi}:\mathcal{T}^N\rightarrow\mathbb{R}$:  
$$\mathcal{L}_\theta^{\pi} = \frac{1}{NH}\sum_{i=1}^N \sum_{t=t_0}^{t_0+H-1} \gamma^{t-t_0} R(s_t^i, a_t^i) + \gamma^H V(s^i_{t_0+T}),$$  
where $N$ is the number of sampled trajectories, $H \ll T$ is a time horizon, $s_t^i$ is the state at time $t$ for trajectory $i$, $a_t^i$ is the action taken at time $t$ for trajectory $i$, and $V(s): \mathcal{S} \rightarrow \mathbb{R}$ is the value function estimating the expected return starting from state $s$. Additionally, SHAC assumes that the policy, reward, and transition functions are differentiable, i.e., $\pi, \hat{F}, \hat{R} \in \mathcal{C}^1$.

The value function is trained to approximate a temporal difference-$\lambda$ formulation \cite{Sutton98} $\hat{V}$, which in SHAC translates to minimize the loss $\mathcal{L}_\theta^{V}:\mathcal{T}^N\rightarrow\mathbb{R}$.

$$ \frac{1}{N}\sum_i^N\left|\left| V(s^i) - \hat{V}(s^i) \right|\right|^2 $$
$$ \hat{V}(s_t) = (1-\lambda) \left(\sum_{k=1}^{h-t-1}\lambda^{k-1}G_t^k\right) + \lambda^{h-t-1}G_t^{h-t}$$
$$ G_t^k = \left(\sum_{l=0}^{k-1}\gamma^l r_{t+l}\right) + \gamma^k V(s_{t+k})$$

$G^t_k$ represents the discounted sum of rewards up to step $k$, plus the bootstrapped value of the state at step $t+k$. The estimates $\hat{V}$ is computed by discounting longer (higher $k$) discounted sums, $G^k_t$.

>>>>

\section{Background \& Notation}\label{sect:background}

\subsection{Markov Decision Processes}
To provide context for our work, we first introduce the reinforcement learning (RL) framework. Each task is modeled as a finite-horizon Markov Decision Process (MDP) $\mathcal{M} = (\mathcal{S}, \mathcal{A}, F, R, T, \gamma, \mu)$, where $\mathcal{S}$ represents the state space, $\mathcal{A}$ denotes the action space, $F:\mathcal{S} \times \mathcal{A} \rightarrow \mathcal{S}$ is the transition function, $R:\mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow \mathbb{R}$ is the reward function, $\gamma \in [0, 1]$ is the discount factor, and $T \in \mathbb{N}$ is the time horizon. The initial state distribution is denoted by $\mu$.

Further, let $\mathcal{T}$ denote the space of trajectories, defined as $\mathcal{T} = (\mathcal{S} \times \mathcal{A} \times \mathcal{S} \times \mathbb{R})^*$, where each trajectory consists of a sequence of starting states, actions, subsequent states, and rewards. We will use the symbols $\tau=((s_0, a_0, s_1, r_0), (s_1, a_1, s_2, r_1), \ldots)$ to denote a generic trajectory in $\mathcal{T}$. When dealing with multiple trajectories, we will use the notation $s_i^\tau, a_i^\tau$, and $r_i^\tau$.

A policy is a function $\pi:\mathcal{S}\rightarrow\mathcal{A}$ that maps states to actions. We can use the policy to sample trajectories from the MDP by iteratively applying the transition and reward function and the policy. In this case, we have that the $i$-th element of a trajectory is $(s_i, a_i, s_{i+1}, r_i) = (s_i, \pi(s_i), F(s_i, a_i), R(s_i,a_i,F(s_i, a_i))$. The initial state $s_0$ is sampled from the initial state distribution $\mu$.

For brevity, we will simply denote with $\tau\sim\pi$ a trajectory sampled by following the policy $\pi$.

The goal of the MDP is to find a policy that maximizes the expected return 
$$\mathbb{E}_{\tau\sim\pi}\left[\sum_{t=0}^{T-1}\gamma^t r_t^\tau \right]$$

\subsection{Neural Networks}
A neural network is a function $f_{\theta}:\mathbb{R}^n\rightarrow\mathbb{R}^m$ parameterized by $\theta$ that maps an input vector $x\in\mathbb{R}^n$ to an output vector $y\in\mathbb{R}^m$. The parameters $\theta$ of the network are learned by minimizing a loss function $\mathcal{L}_\theta$ with respect to the training data.

In this work, we will often approximate functions, such as transition ($F$) or reward ($R$), with neural network. Therefore, for each function $g$, we will denote the corresponding neural network with $g_\theta$. For instance, the transition and its corresponding neural network are denoted as $F$ and $F_{\theta}$, respectively. The corresponding loss will be denoted as $\mathcal{L}_\theta^g$ and it will aim to minimize the differences between $g$ and $g_\theta$.

While we use a unique set of parameters $\theta$ for all the networks. Each network can either use its own subset of $\theta$ or it can share parameters with other networks. In practice, the parameters will never be shared.

@@

To solve an MDP $\mathcal{M}$, SHAC trains a policy $\pi_\theta$ by minimizing the loss function $\mathcal{L}_\theta^{\pi}:\mathcal{T}^N\rightarrow\mathbb{R}$ for a batch of $N$ trajectories $B = \{\tau_1, \ldots, \tau_N\}$ sampled from the MDP. The loss is defined as:  
\begin{equation}\label{eq:shac}
    \mathcal{L}_\theta^{\pi}(B) = \frac{1}{NH}\sum_{\tau\in B} \sum_{t=t_0}^{t_0+H-1} \gamma^{t-t_0} R(s_t^\tau, a_t^\tau) + \gamma^H V_\theta(s^\tau_{t_0+T})
\end{equation}
where $N$ is the number of sampled trajectories, $H \ll T$ is a time horizon, $s_t^\tau$ is the state at time $t$ for trajectory $\tau$, $a_t^\tau$ is the action taken at time $t$ for trajectory $\tau$, and $V(s): \mathcal{S} \rightarrow \mathbb{R}$ is the value function estimating the expected return starting from state $s$. From Equation~\ref{eq:shac}, it is apparent that SHAC assumes that the policy, reward, value, and transition functions are all differentiable, i.e., $\pi_\theta,V_\theta,F,R \in \mathcal{C}^1$. While this is an common requirement for policy and value functions, differentiable transition and reward functions are rarely available.

The value function $V_\theta$ is trained to approximate a temporal difference-$\lambda$ formulation \cite{Sutton98} $V$, which in SHAC translates to minimize the loss $\mathcal{L}_\theta^{V}:\mathcal{T}^N\rightarrow\mathbb{R}$.

$$ \mathcal{L}_\theta^V(B)=\frac{1}{NT}\sum_{\tau\in B}\sum_{t=0}^T\left|\left| V_\theta(s^\tau_t) - V(s^\tau_t) \right|\right|^2 $$
$$ V(s_t) = (1-\lambda) \left(\sum_{k=1}^{h-t-1}\lambda^{k-1}G_t^k\right) + \lambda^{h-t-1}G_t^{h-t}$$
$$ G_t^k = \left(\sum_{l=0}^{k-1}\gamma^l r_{t+l}\right) + \gamma^k V(s_{t+k})$$

$G^t_k$ represents the discounted sum of rewards up to step $k$, plus the bootstrapped value of the state at step $t+k$. The estimates $V$ is computed by discounting longer (higher $k$) discounted sums, $G^k_t$.

### 2025-01-13 paper/icml25/first/sects/experiments.tex
minor + figs

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-1-mlp.pdf}
        \caption{Dispersion, MLP, 1 agent}
        \label{fig:dispersion-mlp-1}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-3-transformer.pdf}
        \caption{Dispersion, Transformer, 3 agents}
        \label{fig:dispersion-transformer-3}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-5-transformer.pdf}
        \caption{Dispersion, Transformer, 5 agents}
        \label{fig:dispersion-transformer-5}
    \end{subfigure}

@@

@@

\paragraph{Scenarios}
As mentioned previously, we choose a few multi-agent scenarios from the VMAS\footnote{\url{https://github.com/proroklab/VectorizedMultiAgentSimulator}} package to test \fname{} aganist SHAC and PPO. In particular, we choose the following scenario:
\begin{compactitem}
    \item \textbf{Dispersion}: a scenario where $n$ agents need to reach $n$ goal locations, in such case the agent receives a sparse reward of $1$. Agents observe only their own position, velocity and relative position to each goal. Despite the transition function being differentiable, the reward function is not. Therefore, only PPO and \fname{} are applicable to this scenario.
    \item \textbf{Transport}: a scenario where $n$ agents need to collaborate to push a package into a goal location. Agents observe their own position, velocity, relative position to the package and relative position between package and the goal. The agents receive a reward proportional to the distance between package and the goal. In this scenario, both the reward and transition functions are differentiable.
    \item \textbf{Sampling}: a scenario where $n$ agents need to collect rewards from a grid. The reward are sampled from $k$ gaussians. Agents observe their own position, velocity, lidar values to since each other, and rewards in a 3x3 grid around themselves.  In this scenario, both the reward and transition functions are differentiable.
    \item \textbf{Discovery}: a scenario where $n$ agents need to collect $k$ goal place randomly. A goal is collected if at least $s$ agents stand in proximity to it, in this case they receive a reward of $1$. If agents collide with each other, they receive a penalty. Every agent observes its position, velocity, lidar measurements to other agents and goals. In this scenario, only the transition function is differentiable. Therefore, only PPO and \fname{} are applicable to this scenario.
\end{compactitem}

@@

\paragraph{Sampling}
TODO

\paragraph{Dispersion}
TODO

>>>>

\begin{figure*}[!t]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-1-mlp.pdf}
        \caption{Dispersion, MLP, 1 agent}
        \label{fig:dispersion-mlp-1}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-3-transformer.pdf}
        \caption{Dispersion, Transformer, 3 agents}
        \label{fig:dispersion-transformer-3}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion-5-transformer.pdf}
        \caption{Dispersion, Transformer, 5 agents}
        \label{fig:dispersion-transformer-5}
    \end{subfigure}

@@

    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/discovery-1-mlp.pdf}
        \caption{Discovery, MLP, 1 agent}
        \label{fig:discovery-mlp-1}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/discovery-3-transformer.pdf}
        \caption{Discovery, Transformer, 3 agents}
        \label{fig:discovery-transformer-3}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{figs/discovery-5-transformer.pdf}
        \caption{Discovery, Transformer, 5 agents}
        \label{fig:discovery-transformer-5}
    \end{subfigure}

    \caption{Comparison between \fname{}, PPO, and SHAC for increasing number of agents for Dispersion, Transport, Sampling, and Discovery scenarios.}
    \label{fig:experiments}

@@

\paragraph{Scenarios}
As mentioned previously, we choose a few multi-agent scenarios from the VMAS\footnote{\url{https://github.com/proroklab/VectorizedMultiAgentSimulator}} package to test \fname{} aganist SHAC and PPO. In particular, we choose the following scenario:
\begin{compactitem}
    \item \textbf{Dispersion} (Fig.~\ref{fig:dispersion}): a scenario where $n$ agents need to reach $n$ goal locations, in such case the agent receives a sparse reward of $1$. Agents observe only their own position, velocity and relative position to each goal. Despite the transition function being differentiable, the reward function is not. Therefore, only PPO and \fname{} are applicable to this scenario. The maximum reward is $n$.
    \item \textbf{Transport} (Fig.~\ref{fig:transport}): a scenario where $n$ agents need to collaborate to push a package into a goal location. Agents observe their own position, velocity, relative position to the package and relative position between package and the goal. The agents receive a reward proportional to the distance between package and the goal. In this scenario, both the reward and transition functions are differentiable. The maximum reward varies from environement to environment. It is the scaled distance between the package and the goal times the number of agents.
    \item \textbf{Discovery} (Fig.~\ref{fig:discovery}): a scenario where $n$ agents need to collect $k$ goal place randomly. A goal is collected if at least $s$ agents stand in proximity to it, in this case they receive a reward of $1$. If agents collide with each other, they receive a penalty. Every agent observes its position, velocity, lidar measurements to other agents and goals. In this scenario, only the transition function is differentiable. Therefore, only PPO and \fname{} are applicable to this scenario. When $n=1$ we set $s=1$ and $k=7$. When $n>1$ we set $s=2$. The maximum rewards depends from the time available, but we set it to $k$.
    \item \textbf{Sampling} (Fig.~\ref{fig:sampling}): a scenario where $n$ agents need to collect rewards from a grid. The reward are sampled from $k$ gaussians. Agents observe their own position, velocity, lidar values to each other, and rewards in a 3x3 grid around themselves.  In this scenario, both the reward and transition functions are differentiable. The maximum reward depends on where the gaussians are placed.
\end{compactitem}

@@

\paragraph{Discovery}
Similar consideration hold also for the discovery scenario. In the single agent setting, shown in Fig.~\ref{fig:discovery-mlp-1}, \fname{} performs better, but comparitavely, to PPO . However, with $3$ and $5$ agents, shown in Fig.~\ref{fig:discovery-transformer-3} and \ref{fig:discovery-transformer-5}, PPO fails to converge, meanwhile \fname{} performs better.

\paragraph{Sampling}

### 2025-01-13 paper/icml25/first/sects/shacpp.tex
minor + figs

@@

As mentioned previously, we aim to address the limitations of SHAC by approximating the gradients of a differentiable simulator with neural networks. In particular, we aim to train a neural network to approximate the reward function by minimizing: 
$$ \mathbb{E}_{s,a}\left[\left|\left| R_{\theta}(s, a) - \hat{R}(s, a) \right|\right|^2\right] $$

In practice, this is achieved by leveraging a \emph{replay buffer} to cache the state-action-reward triplets obtained from the environment. The neural network is then trained to minimize the MSE loss. Specifically, we minimize the loss $\mathcal{L}_\theta^{R}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \frac{1}{\left|B\right|}\sum_{s,a,r \in B}\left|\left| R_\theta(s, a) - \underbrace{r}_{\hat{R}(s, a)} \right|\right|^2 $$
Where $B$ represents a batch of state-action-reward triplets sampled uniformly from the cache.

Similarly, we aim to train a neural network to approximate the transition function. However, to avoid vanishing or exploding gradients, we employ a non-recursive approach similar similar to Action World Model~\cite{Ma24}. This corresponds to train a neural network that given a starting observation from $\mathcal{S}$ and a sequence of $N$ actions from $\mathcal{A}^H$, predicts the next $N$ observations from $\mathcal{S}^H$. In practice, we aim to minimize:

$$ \mathbb{E}_{\tau}\left[\sum_{i=1}^H \left|\left| F_{\theta}(s_0, a_0, \dots, a_N)_i - \hat{F}(s_i, a_i) \right|\right|^2\right] $$

Where $\tau\in\mathcal{T}$. In practice, this is achieved by caching trajectories and then training the neural network to minimize the MSE loss. That is, we minimize the loss $\mathcal{L}_\theta^{F}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \frac{1}{NH}\sum_{\tau \in B}\sum_{i=1}^H\left|\left| F_{\theta}(s_0, a_0, \dots, a_H)_i - \underbrace{s_{i}}_{\hat{F}(s_{i-1}, a_{i-1})} \right|\right|^2 $$
Where $B$ represents a batch of trajectories sampled uniformly from the cache.

Further, we modify the SHAC policy loss, $\mathcal{L}_\theta^{\pi}$, to include a term that penalizes actions outside the action space. That is, suppose the action space is $(a,b)$ (with $a<b$), we minimize the loss $\mathcal{L'}_\theta^\pi:$

$$\mathcal{L}_\theta^\pi + \alpha \frac{1}{NH}\sum_{i=0}^N\sum_{t=t_0}^{t_0+H-1} d(a_t^i,(a,b))$$
$$d(x,(a,b)) = \begin{cases}(a-x)^2 & \text{if } x < a \\ (x-b)^2 & \text{if } x > b \\ 0 & \text{otherwise} \end{cases}$$

@@

To summarize, our algorithm, aims to minimize the 4 presented losses, $\mathcal{L'}_\theta^{\pi}$, $\mathcal{L}_\theta^V$, $\mathcal{L}_\theta^R$, and $\mathcal{L}_\theta^F$. The parameters $\theta$ can be either shared or separated among the different networks, however, we always keep them separated in our experiments. A minimal pseudocode of our algorithm is presented in \ref{alg:shacpp}. Figure~\ref{fig:shacpp} presents an unrolled view of the \fname{} algorithm.

@@

        \STATE $\theta \gets \theta - \eta_\pi \nabla \mathcal{L'}^\pi_\theta(\hat{r},\hat{v},\gamma,\alpha)$
        \STATE $\theta \gets \theta - \eta_V \nabla \mathcal{L}^V_\theta(\hat{v},s,v)$
        \STATE $\theta \gets \theta - \eta_R \nabla \mathcal{L}^R_\theta(\hat{r},s,r)$
        \STATE $\theta \gets \theta - \eta_F \nabla \mathcal{L}^F_\theta(\hat{s},s,a)$

>>>>

        \STATE $\theta \gets \theta - \eta_\pi \nabla \mathcal{L'}^\pi_\theta(\hat{r},\hat{v},\gamma,\alpha)$
        \STATE $\theta \gets \theta - \eta_V \nabla \mathcal{L}^V_\theta(\hat{v},s,v)$
        \STATE $\theta \gets \theta - \eta_R \nabla \mathcal{L}^R_\theta(\hat{r},s,a,r)$
        \STATE $\theta \gets \theta - \eta_F \nabla \mathcal{L}^F_\theta(\hat{s},s,a)$

@@

As mentioned previously, we aim to address the limitations of SHAC by approximating the gradients of a differentiable environment ($R$ and $F$) with neural networks ($R_\theta$ and $F_\theta$). In particular, we aim to train a neural network to approximate the reward function by minimizing: 
$$ \mathbb{E}_{\tau\sim\pi}\left[\left|\left| \frac{1}{T}\sum_{i=0}^{T-1} R_{\theta}(s_i^\tau, a_i^\tau, s_{i+1}^\tau) - R(s_i^\tau, a_i^\tau, s_{i+1}^\tau) \right|\right|^2\right] $$

In practice, this is achieved by leveraging a \emph{replay buffer} to cache the state-action-reward triplets obtained from the environment. The neural network is then trained to minimize the mean squared error (MSE) loss. Specifically, we minimize the loss $\mathcal{L}_\theta^{R}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \mathcal{L}_\theta^R(B) = \frac{1}{N}\sum_{s,a,r \in B}\left|\left| R_\theta(s, a) - r \right|\right|^2 $$
Where $B$ represents a batch of state-action-reward triplets sampled uniformly from the cache.

Similarly, we aim to train a neural network to approximate the transition function. However, to avoid vanishing or exploding gradients, we employ a non-recursive approach similar similar to Action World Model~\cite{Ma24}. This corresponds to train a neural network that given a starting observation from $\mathcal{S}$ and a sequence of $H$ actions from $\mathcal{A}^H$, predicts the next $H$ observations from $\mathcal{S}^H$. We aim to minimize:

\begin{equation}
\begin{aligned}
\mathbb{E}_{\tau\sim\pi}\left[\frac{1}{NH}\sum_{t=0}^{T-H-1}\sum_{i=0}^{H-1} \left|\left| F_{\theta}(s_t^\tau, a_t^\tau, \dots, a_{t+H-1}^\tau)_i - \right.\right.\right. \\ \left.\left.\left. s_{t+i+1}^\tau \right|\right|^2\right] 
\end{aligned}
\end{equation}

In practice, this is achieved by caching trajectories and then training the neural network to minimize the MSE loss. That is, we minimize the loss $\mathcal{L}_\theta^{F}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \frac{1}{NH}\sum_{\tau \in B}\sum_{t=t_0}^{t_0+H-1}\left|\left| F_{\theta}(s_t^\tau, a_t^\tau, \dots, a_{t+H-1}^\tau)_i - s_{t+1}^\tau \right|\right|^2 $$
Where $B$ represents a batch of trajectories with horizon $H$ sampled uniformly from the cache.

Further, we modify the SHAC policy loss, $\mathcal{L}_\theta^{\pi}$, to include a term that penalizes actions outside the action space. That is, suppose the action space is $(a,b)$ (with $a<b$), we minimize the loss $\mathcal{L'}_\theta^\pi:$
\begin{figure*}[!t]
    \centering
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion.png}
        \caption{Dispersion}
        \label{fig:dispersion}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/transport.png}
        \caption{Transport}
        \label{fig:transport}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/discovery.png}
        \caption{Discovery}
        \label{fig:discovery}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/sampling.png}
        \caption{Sampling}
        \label{fig:sampling}
    \end{subfigure}
    \caption{VMAS environments used in the experiments.}\label{fig:scenarios}
\end{figure*}

$$\mathcal{L}_\theta^\pi(B) + \alpha \frac{1}{NH}\sum_{\tau\in B}\sum_{t=t_0}^{t_0+H-1} d(a_t^\tau,(a,b))$$
$$d(x,(a,b)) = \begin{cases}(a-x)^2 & \text{if } x < a \\ (x-b)^2 & \text{if } x > b \\ 0 & \text{otherwise} \end{cases}$$

@@

To summarize, our algorithm, aims to minimize the 4 presented losses, $\mathcal{L'}_\theta^{\pi}$, $\mathcal{L}_\theta^V$, $\mathcal{L}_\theta^R$, and $\mathcal{L}_\theta^F$. A minimal pseudocode of our algorithm is presented in \ref{alg:shacpp}. Figure~\ref{fig:shacpp} presents an unrolled view of the \fname{} algorithm.

@@
